---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icecream

<!-- badges: start -->
<!-- badges: end -->

Ported from [gruns/icecream](https://github.com/gruns/icecream).

## Installation

You can install from GitHub with:

``` r
devtools::install_github("lewinfox/icecream")
```

## Inspect variables

The `ic()` function prints its argument and its value. It also returns the value of the evaluated
argument.

```{r example}
library(icecream)

is_negative <- function(x) x < 0

ic(is_negative(1))

ic(is_negative(-1))
```

More complex inputs are summarised to avoid cluttering the terminal.

```{r}
df <- ic(iris)

my_list <- ic(list(a = 1, b = 3, c = 1:100))
```

## Inspect execution

Calling `ic()` with no arguments causes it to print out the file, line and parent function it was
called from.

``` r
# demo.R
f1 <- function(x) {
  ic()
  if (x > 0) {
    f2()
  }
}

f2 <- function() {
  ic()
}

```

``` r
f1(-1)
#> ℹ ic| demo.R:2 in `f1()

f1(1)
#> ℹ ic| demo.R:2 in `f1()`
#> ℹ ic| demo.R:9 in `f2()`
```

In the case of functions that haven't been `source()`d or loaded from a package, the function's
environment will be displayed.

``` r
orphan_func <- function() {
  ic()
  TRUE
}

orphan_func()
#> ℹ ic| <environment: R_GlobalEnv>: `orphan_func()`
```

## Enable / disable

The `ic_enable()` and `ic_disable()` functions enable or disable the `ic()` function. If disabled,
`ic()` will return the result of evaluating its input but will not print anything.

```{r}
ic_enable()

ic(mean(1:100))

ic_disable()

ic(mean(1:100))
```

## TODO:

* Implement `ic.format()` (see [here](https://github.com/gruns/icecream#miscellaneous)).
